Place of death in NRS COVID-19 deaths: visualisation for 2022 data

### 2022-03-28 covid_nrs_place_of_death_visualisations_2022.Rmd
added bits of code to keep 2015-2019 historic period in 2022, and a place of death figure using the same

# removed_other_category_text <- glue::glue("Excluding 'Other' institutional place of death, N={merged_deaths_overall %>% filter(sex == 'all' & age == 'all' & place_of_death == 'Other') %>% .$deaths_all_causes %>% sum} for weeks 1-{max(merged_deaths_overall %>% filter(year==2020) %>% .$week_number)} in 2020 and 1-{max(merged_deaths_overall %>% filter(year==2021) %>% .$week_number)} in 2021.")
removed_other_category_text <- glue::glue("Excluding 'Other' institutional place of death.")
```

@@

```{r}
figure_place_of_death_home_only <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%  # home death only
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,removed_other_category_text,source_nrs_text), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom="text",x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 50, label = "Historic range (min & max)", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 29, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 33, year_number = 2020), y = 65, yend = 210, curvature = 0.20, colour = "#4477aa", alpha = 0.5, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom="text", x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 100, label = "Historic mean", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 22, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 25, year_number = 2020), y = 115, yend = 270, curvature = 0.20, colour = "grey20", linetype="dashed", arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "'22", hjust = 0, colour = "grey40") +
  NULL

@@

>>>>

# removed_other_category_text <- glue::glue("Excluding 'Other' institutional place of death, N={merged_deaths_overall %>% filter(sex == 'all' & age == 'all' & place_of_death == 'Other') %>% .$deaths_all_causes %>% sum} for weeks 1-{max(merged_deaths_overall %>% filter(year==2020) %>% .$week_number)} in 2020 and 1-{max(merged_deaths_overall %>% filter(year==2021) %>% .$week_number)} in 2021.")
removed_other_category_text <- glue::glue("Excluding 'Other' institutional place of death.")

historic_period_2015_2019_text <- "Historic mean & range for 2015-2019 throughout."
historic_period_2022_text <- "2020-21 use 2015-2019 historic mean & range; 2022 uses 2016-2019 & 2021."
```

@@

```{r}
figure_place_of_death_home_only <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%  # home death only
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,historic_period_2022_text,source_nrs_text), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom="text",x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 50, label = "Historic range (min & max)", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 29, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 33, year_number = 2020), y = 65, yend = 210, curvature = 0.20, colour = "#4477aa", alpha = 0.5, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom="text", x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 100, label = "Historic mean", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 22, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 25, year_number = 2020), y = 115, yend = 270, curvature = 0.20, colour = "grey20", linetype="dashed", arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "'22", hjust = 0, colour = "grey40") +
  NULL

@@

### Home deaths with 2015-2019 historic period means & range only

```{r}
merged_deaths_overall_2015_2019_historic_period <-
  bind_rows(
    ## pre-2022 data
    merged_deaths_overall %>% filter(year < 2022),
    ## 2022 data
    merged_deaths_overall %>%
      filter(year >= 2022) %>%  # 2022 or later!?
      select(-matches("deaths\\_mean|deaths\\_min|deaths\\_max")) %>%  # remove data using NRS historic period
      mutate(historic_period = "2015-2019") %>%  # join the previous hsitoric period data instead
      left_join(historical_weekly_average_overall, by = c("historic_period","week_number","place_of_death")) %>%
      compute_moving_average_for_current_data(  # not strictly required here but will be easier to apply in other places if included
        current_data = ., 
        historical_data = 
          weekly_deaths_2015_2019_overall %>% mutate(year=as.integer(year)) %>% drop_na(year),  # remove the entries with 'average' for year
        num_weeks_to_average = 4)
  )

figure_place_of_death_home_2015_2019_historic_period <-
  merged_deaths_overall_2015_2019_historic_period %>%
  filter(sex == "all" & age == "all") %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%  # home death only
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,historic_period_2015_2019_text,source_nrs_text), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom="text",x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 50, label = "Historic range (min & max)", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 29, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 33, year_number = 2020), y = 65, yend = 210, curvature = 0.20, colour = "#4477aa", alpha = 0.5, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom="text", x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 100, label = "Historic mean", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 22, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 25, year_number = 2020), y = 115, yend = 270, curvature = 0.20, colour = "grey20", linetype="dashed", arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "2022", hjust = 0, colour = "grey40") +
  NULL

figure_place_of_death_home_2015_2019_historic_period %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death_home_2015-2019_range", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 18, height = 14, dpi = 300, units = "cm")
```
